---
title: "Data_Wrangling_Ex2"
author: "Blae Quayle"
date: "27 April 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Data Wrangling Exercise 2

## Step 0: Load the data into R studio and relevant packages

```{r message=FALSE, warning=FALSE}
library(readr)
library(dplyr)
library(tidyr)

titanic <- read.csv("~/Documents/Work/Data Science/Data_Wrangling_Ex2/titanic3.csv")
data.frame(titanic)
tbl_df(titanic)
```

## Step 1: Find missing values and replace them with S in embark variable

```{r}
titanic_embarked <- titanic %>%
  mutate(embarked = sub("^$", "S", embarked))
```

## Step 2: Find mean of age column and use to populate missing values

```{r}
mean_age <- mean(titanic$age, na.rm = TRUE)
print(mean_age)

titanic_ages <- titanic_embarked %>%
  replace_na(list(age = mean_age))
```

The mean of the ages is `r mean_age`
The median of the ages is `r median_age`

The median is probably more appropriate here as the ages are skewed to the left ie. younger.
Using the mean gives equal weight to the oldest people, who make up a smaller proportion.

## Step 3: Fill empty slots in boat column with a dummy value

```{r}
titanic_lifeboats <- titanic_ages %>%
  mutate(boat = sub("^$", "None", boat))  
```

## Step 4: What does a missing cabin value mean and does it make sense to fill it in?

Missing values could be filled with 'Unknown' but it wouldn't make the dataset clearer.
Missing values could indicate non-survival, therefore no record of their cabin number.

### Create a new column has_cabin_number which has 1 if there is a cabin numnber and 0 otherwise

```{r}
titanic_cabin <- titanic_lifeboats %>%
  mutate(has_cabin_number = ifelse(cabin == "", 0, 1))
```

The majority of people who died do not have a cabin number stated.

## Step 5: Save new data file as .csv

```{r}
write.csv(titanic_cabin, "titanic_clean.csv")
```

