---
title: "Data Wrangling Ex1"
author: "Blae Quayle"
date: "26 April 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Data Wrangling Excercise 1

Clean the dataset provided to make it easier to analyse and visualise. Note that new dataframes have been created at each stage to track the separate stages and that this is only feasible if the dataset is small.

## Step 0: Load the data into R studio and relevant packages

```{r message=FALSE, warning=FALSE}
library(readr)
library(dplyr)
library(tidyr)
pp <- read_csv("~/Documents/Work/Data Science/refine_original.csv")
tbl_df(pp)
```

## Step 1: Clean up brand names to remove mispellings

```{r}
clean_companies <- pp %>%
  mutate(company = tolower(company)) %>% 
  mutate(company = case_when(
    .$company %in% c('phillips', 'phillps', 'phllips', 'fillips', 'phlips') ~ 'philips',
    .$company %in% c('akz0', 'ak zo') ~ 'akzo',
    .$company %in% c('unilvr', 'unlever', 'unilver') ~ 'unilever',
    TRUE ~ .$company)
  )
```

## Step 2: Separate the product code and number

```{r}
separate_product <- clean_companies %>%
  separate(`Product code / number`, c("product_code", "product_number"), sep = "-", remove = TRUE)
```

## Step 3: Add product categories

```{r}
product_categories <- separate_product %>%
  mutate(category = case_when( 
    .$product_code %in% c('p') ~ 'Smartphone',
    .$product_code %in% c('v') ~ 'TV',
    .$product_code %in% c('x') ~ 'Laptop',
    .$product_code %in% c('q') ~ 'Tablet',
    TRUE ~ .$product_code ))
```

## Step 4: Add the full address for geocoding

```{r}
combined_address <- product_categories %>%
  unite(full_address, address:country, sep = ", ", remove = TRUE)
```

## Step 5: Create dummy variables for company and product category

```{r}
dummy_variables <- combined_address %>%
  mutate(company_binary = 1, product_binary = 1) %>%
  mutate(company = paste0("company_", company), category = paste0("product_", category)) %>%
  spread(company, company_binary, fill = 0) %>%
  spread(category, product_binary, fill = 0)

```

## Step 6: Save the new dataframe as a .csv file and submit to GitHub

```{r}
write.csv(dummy_variables, "refine_clean.csv")
```

